---
title: "Bike Share Demand & Weather Analytics"
output: html_document
---

```{r setup}
df <- read.csv("data/bikeshare.csv")
df$weathersit <- factor(df$weathersit)
summary(df[, c("cnt","weathersit","temp","hum","windspeed")])
```

## Demand distribution and weather groups
```{r}
hist(df$cnt, breaks=40, main="Distribution of Bike Rentals", xlab="Rentals")
boxplot(cnt ~ weathersit, data=df, ylab="Bike rentals", xlab="Weather situation")
```

## Multiple regression
```{r}
base_model <- lm(cnt ~ weathersit + temp + hum + windspeed, data=df)
summary(base_model)
```

## Interaction model
```{r}
interaction_model <- lm(cnt ~ weathersit * (temp + hum + windspeed), data=df)
summary(interaction_model)
```

The interaction terms test whether slopes for temperature, humidity, or windspeed differ by weather condition.

## Translate normalized temperature to a 10°C change
```{r}
delta_temp_norm <- 10 / 41
coefs <- coef(interaction_model)

temp_effect_for_level <- function(level_number) {
  base <- coefs[["temp"]]
  if (level_number == 1) return(base * delta_temp_norm)
  pattern <- paste0("weathersit", level_number, ":temp|temp:weathersit", level_number)
  interaction <- unname(coefs[grep(pattern, names(coefs))])
  (base + interaction) * delta_temp_norm
}

temp_effect_for_level(1)
if ("2" %in% levels(df$weathersit)) temp_effect_for_level(2)
if ("3" %in% levels(df$weathersit)) temp_effect_for_level(3)
```

## Log-demand model
```{r}
df$log_cnt <- log(df$cnt)
log_model <- lm(log_cnt ~ weathersit * (temp + hum + windspeed), data=df)
summary(log_model)

# For a coefficient b and change delta, exp(b*delta) is the multiplicative effect.
exp(coef(log_model)[["temp"]] * delta_temp_norm)
```

## Example prediction
```{r}
new_case <- data.frame(
  weathersit=factor(levels(df$weathersit)[1], levels=levels(df$weathersit)),
  temp=25/41, hum=0.5, windspeed=5/67
)
predict(interaction_model, newdata=new_case)
exp(predict(log_model, newdata=new_case))
```

## High-demand logistic model
The source coursework defined a high-demand day as more than 7,000 rentals.

```{r}
df$busy <- as.integer(df$cnt > 7000)
busy_model <- glm(
  busy ~ weathersit * (temp + hum + windspeed),
  data=df, family=binomial
)
summary(busy_model)

# Temperature odds ratio for a 10°C change in the reference weather condition
exp(coef(busy_model)[["temp"]] * delta_temp_norm)
```

## Interpretation
Interaction coefficients should not be interpreted in isolation. For a non-reference weather group, the relevant temperature slope is the reference temperature coefficient plus that group's temperature interaction coefficient.
